

### More Data

Now let's look at the association between illegal immigration and crime rates from 1990-2014.

We'll start by analyzing violent crime rates.

```{r}

# datasets needed and names assigned

uscrime <- read.csv("USCrime.csv")

illegal <- read.csv("illegal.csv")

usapop <- read.csv("usapop.csv")

usapop <- setNames(usapop, c("date", "uspop"))

uscrime <- setNames(uscrime, c("year", "population", "violent", "property", "violent_percentage", "pre"))

illegal <- setNames(illegal, c("illegal_year", "illegal_pop"))

# the following makes it so that we don't have abbreviated numbers showing in the x and y axes.

options(scipen = 999)

# I now have to shorten the vector because we're only looking at 1990-2014 now
# The name below refers to 19-ninety, the year we begin

ninetyviolentcrime <- c(uscrime$violent[31:55])


violentillegal <- (lm(illegal$illegal_pop ~ ninetyviolentcrime))

plot(ninetyviolentcrime, illegal$illegal_pop, xlab ="Violent crime incidents", ylab ="Immigrants in a year")

abline(violentillegal, col="blue")

summary(violentillegal)

cor(ninetyviolentcrime, illegal$illegal_pop)

```

Wow! The correlation between illegal immigration and violent crime is -0.90. That's a very strong relationship, and this time the relationship is linear and negative. The p-value is a miniscule 0.00000000062. 

What happens when we adjust for illegal immigration's impact on the greater US population?

```{r}

illegal_adjusted <- c(illegal$illegal_pop)/c(usapop$uspop[31:55])

violentillegal_adjusted <- (lm(illegal_adjusted ~ ninetyviolentcrime))

plot(ninetyviolentcrime, illegal_adjusted, xlab ="Violent crime incidents", ylab ="Immigrants as proportion")

abline(violentillegal_adjusted, col="blue")

cor(ninetyviolentcrime, illegal_adjusted)

summary(violentillegal_adjusted)

```

There's still a negative, linear relationship here. The strength is strong. The correlation didn't change much. It just dipped slightly to -0.88, and the p-value is still a tiny 0.0000000067.

What happens when we look at illegal immigration and property crime?

```{r}

ninetypropertycrime <- c(uscrime$property[31:55])


propertyillegal <- (lm(illegal$illegal_pop ~ ninetypropertycrime))

plot(ninetypropertycrime, illegal$illegal_pop, xlab ="Property crime incidents", ylab ="Immigrants in a year")

abline(propertyillegal, col="blue")

summary(propertyillegal)

cor(ninetypropertycrime, illegal$illegal_pop)

```

Interesting. As it relates to illegal immigration, the property crime data looks very similar to the violent crime data. There's a strong, negative, linear relationship. The correlation is -0.90. The p-value is 0.0000000012.

Let's see what happens when we look at illegal immigration as it relates to property crime, this time adjusting for population impact.

```{r}

propertyillegal_adjusted <- (lm(illegal_adjusted ~ ninetypropertycrime))

plot(ninetypropertycrime, illegal_adjusted, xlab ="Preoperty crime incidents", ylab ="Immigrants as proportion")

abline(propertyillegal_adjusted, col="blue")

cor(ninetypropertycrime, illegal_adjusted)

summary(propertyillegal_adjusted)


```

That adjustment didn't have much of an impact. The attributes of the relationship remain the same. The correlation lessens slightly but is still a strong -0.87, with a p-value of 0.000000023.